Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-02-28 WSSSPE1_report/WSSSPE1_summary_report.tex
merge, todo: smoothen and expand

For the purpose of studying software case studies, we classify the software
projects discussed in the workshop in three broad categories. The first is 
utility software, software tools that enable and/or facilitate the
development of other tools and techniques to carry out scientific work. The second is
scientific software, software that was developed with an aim to solve a
class of scientific problems. The third is infrastructure software, software
developed to efficiently utilize new research infrastructure.

@@

%and point it out in our
%discussion.

@@

@@

\subsection{Infrastructure software}
%TODO: expand
Infrastructure software is developed to better utilize particular, often new,
infrastructures and architectures, e.g., \emph{MVAPICH}, \emph{VisIT}. Explores
uncharted territories. Highest risk since the projects are experimental.
\note{Dan: so is the only difference here that this category is aimed at new
systems, where utility software is not?  If so, perhaps this should be
incorporated into that subsection, so there would just be two subsections.
Something in the framing of this subsection doesn't feel right to me.
FL: I respectfully disagree the the only reason to develop frameworks is
to better utilize (new or old) infrastructure. Another really important reason
is to modularize development, making it much easier for an individual developer/user
to integrate their own code.}
Successful projects reap high rewards and have longer usage span. By nature,
they overlap with the utility software as they must serve as a facilitator to
run ordinary codes. Long requirement-development-test cycles.  Wide community
support but tends to be handled by specialists rather then end users. Promotes
collaborations across the breadth (e.g., different departments) and depth
(e.g., stakeholders within a department) of community, one of the key
ingredient of a sustainable process. Ferenbaugh~\cite{Ferenbaugh_WSSSPE}
discusses the risks and challenges associated with software projects dealing
with new architectures. To leverage the power of accelerators such as GPUs and
MICs new code and libraries are required. The experience of efforts as
described in~\cite{Ferenbaugh_WSSSPE} met with a limited success but
nonetheless with many invaluable lessons were learned about influential
cultural and technical aspects in sustainable software development practices.

>>>>

For the purpose of studying software case studies, we classify the software
projects discussed in the workshop in two broad categories. The first is
utility software, software tools that enable and/or facilitate the development
of other tools and techniques to carry out scientific work. This includes the
software developed to efficiently utilize new research infrastructures. The
second is scientific software, software that was developed with an aim to solve
a class of scientific problems.

@@

@@

%\subsection{Infrastructure software}
\todo{expand}
Utility software is often developed aimed at better utilizing particular, new
infrastructures and architectures, e.g., \emph{MVAPICH}, \emph{VisIT}.
Development in many cases is often a high risk/reward undertaking and explores
uncharted territories.

%
%\note{Dan: so is the only difference here that this category is aimed at new systems, where utility software is not?  If so, perhaps this should be incorporated into that subsection, so there would just be two subsections.  Something in the framing of this subsection doesn't feel right to me.  FL: I respectfully disagree the the only reason to develop frameworks is to better utilize (new or old) infrastructure. Another really important reason is to modularize development, making it much easier for an individual developer/user to integrate their own code. Ketan: I think in this section I wanted to bring up a class of software projects which are undertaken with the main aim of tackling new systems and architectures in mind. New infrastructures such as GPU, Cloud, fast IB, federal science networks, etc. were on my mind as I see many submissions around them in the workshop.}
%
Successful projects reap high rewards and have longer usage span. Long
requirement-development-test cycles.  Wide community support but tends to be
handled by specialists rather then end users. Promotes collaborations across
the breadth (e.g., different departments) and depth (e.g., stakeholders within
a department) of community, one of the key ingredient of a sustainable process.
Ferenbaugh~\cite{Ferenbaugh_WSSSPE} discusses the risks and challenges
associated with software projects dealing with new architectures. To leverage
the power of accelerators such as GPUs and MICs new code and libraries are
required. The experience of efforts as described in~\cite{Ferenbaugh_WSSSPE}
met with a limited success but nonetheless with many invaluable lessons were
learned about influential cultural and technical aspects in sustainable
software development practices.

@@
